---
title: "Cake Puzzle"
author: "Eamonn O'Brien"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document:
    toc: true
    toc_depth: 2
  fig_height: 6
  fig_width: 8
header-includes:
- \usepackage{eso-pic,graphicx,transparent}
- \usepackage{graphicx}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \setlength\headheight{22pt}
- \fancyfoot[RO]{New Scientist Puzzle}
- \usepackage{lastpage}
- \cfoot{Page \thepage\ of \pageref{LastPage}}
---

\newpage  
\tableofcontents  
\listoffigures
\listoftables
\newpage

```{r set-options, echo=FALSE, cache=FALSE, warning = FALSE}

 

         set.seed(123)
        startTime<-proc.time()
        library(knitr)
         options(width=60)

         opts_chunk$set(comment = "", warning = FALSE, message = FALSE,
                       echo = TRUE, tidy = TRUE, size="tiny",  cache=FALSE,
                       progress=TRUE, tidy.opts=list(width.cutoff=60),
                         fig.width=7, fig.height=3.5,
                       cache.path = 'program_Cache/',
                       fig.path='figure/')
         
        # opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE) 
         
        knitr::knit_hooks$set(inline = function(x) {
          knitr:::format_sci(x, 'md')
        })
         
        
        options(scipen=999)  # remove scientific notation
        

        
        
 
```

```{r ,echo=FALSE}

p2 <- function(x) {formatC(x, format="f", digits=4)}

```


## Function to implement Monte Carlo simulation 

```{r function}


runSim <- function(nSim=1000) {
  #set up an array to store parameter estimates
  estArray <- array(0, dim=c(nSim,1))
  
  for (i in 1:nSim) {

    A <- runif(1)
    B <- runif(1)
    C <- runif(1)
    
    if ( (B > A && B < C)  | (A > B && A < C) ) {res=1} else {res=0}
   
    estArray[i] <-res 
  }
  list(estArray=estArray )
}


x<- runSim(nSim=99999)
mean(x$estArray, na.rm=TRUE)

 
```


## Analytical solution 

```{r analytical solution}

```

Simulation is a little overkill for this problem, but a good programming excercise. We can use permutations. 

Denote the location of the candles as A and C and the location of the cut B. Remember that repetition is allowed in permutations unlike in combinations.

$\sf{^n}{p}{_r}$ : which means the number of permutations of n items taken r items at a time.

For example; given 3 letters ABC find $\sf{^3}{p}{_3}$

ABC, BAC, BCA, CBA, CAB, ACB

which mean that there are 6 ways, in other words $\sf{^3}{p}{_3}$= 6. 

Lets look at the permutation function using factorials:

n!/(n-r)!  = 3!/(3-3)! = 3x2x1/0! = 6/1 = 6.

We can see that B is in the middle of 'ABC, BAC, BCA, CBA, CAB, ACB' 2 out of 6 times. So the chance that both pieces of the cake have a candle on them is 1/3.


\newpage  

## Computing Environment

```{r}

sessionInfo()

```

```{r echo=FALSE}

stopTime<-proc.time()

```

This took `r (stopTime-startTime)[1][[1]]` seconds to execute.

 